plots with prolactin

### 2020-04-06 analysis/16_PRL.Rmd
tables with prl counts

source("../R/themes.R") 
source("../R/functions.R")
source("../R/icons.R")

@@

```{r data}
hormones <- read_csv("../results/hormones.csv")
hormones$treatment <- factor(hormones$treatment, levels = alllevels)

@@

p1 +
  annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.8) +
  annotation_custom(bldg, ymin = -20, ymax = 0, xmin = -5.8) +
  annotation_custom(lay, ymin = -20, ymax = 0, xmin = -3.8) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -1.8) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 0.6) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 2.4) +
  annotation_custom(hatch, ymin = -20, ymax = 0, xmin = 4.4) +
  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 6.4) +
  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 8.4) + 
  ylim(c(-10,120))

@@

## Prolactin (_PRL_) and _BRCA1_ expression in the pituitary

```{r candidatevsds}
colData.pit <-  read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData.pit <- colData.pit %>% filter(tissue == "pituitary")

vsd.pit <- read.csv("../results/06_pitallvsd.csv", header = T, row.names = 1)

savecols <- as.character(colData.pit$V1) 
savecols <- as.vector(savecols) 
vsd.pit <- vsd.pit %>% dplyr::select(one_of(savecols)) 
vsd.pit$genes <- row.names(vsd.pit)

candidategenes <- c("PRL", "BRCA1", "PRLR")
candidates.pit <- vsd.pit %>% dplyr::filter(genes %in% candidategenes)
candidates.pit <- as.data.frame(candidates.pit)
row.names(candidates.pit) <- candidates.pit$genes
candidates.pit$genes <- NULL
candidates.pit <- as.data.frame(t(candidates.pit))
candidates.pit$V1 <- row.names(candidates.pit) 
candidates.pit <- left_join(colData.pit, candidates.pit) 
candidates.pit$treatment <- factor(candidates.pit$treatment, levels = alllevels)
candidates.pit$study2 <- ifelse(candidates.pit$treatment %in% charlevels, "charcterization",
                              ifelse(candidates.pit$treatment %in% levelstiming, "timing",
                                     ifelse(candidates.pit$treatment %in% levelsremoval, "removal", NA)))
tail(candidates.pit)

@@

p2 <- candidates.pit %>%
  dplyr::filter(treatment %in% charlevels) %>%
  ggplot(aes(x = treatment, y = PRL)) + 
    geom_boxplot(aes(fill = treatment, color = sex)) + 
    theme_B3() +
  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
  scale_color_manual(values = sexcolors) +
  scale_fill_manual(values = colorscharnew) +
  labs(y = "PRL", x = "parental care stages", subtitle = "Pituitary" ) +
  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) +
  theme(axis.title.y  = element_text(face = "italic")) 
p2
p2+
  annotation_custom(control, ymin = 13.5, ymax = 15, xmin = -7.5) +
  annotation_custom(bldg, ymin = 13.5, ymax = 15, xmin = -5.5) +
  annotation_custom(lay, ymin = 13.5, ymax = 15, xmin = -3.5) +
  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = -1.5) +
  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 0.5) +
  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 2.5) +
  annotation_custom(hatch, ymin = 13.5, ymax = 15, xmin = 4.5) +
  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 6.5) +
  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 8.5)

```{r hormonesandgenes}
plot_grid(p1 + theme(legend.position ="none",
                     axis.title.x = element_blank(),
                     axis.text.x = element_blank()),
          p2 + theme(legend.position = c(0.85,0.3)) +
            labs(subtitle = NULL) +
  annotation_custom(control, ymin = 13.5, ymax = 15.5, xmin = -7.5) +
  annotation_custom(bldg, ymin = 13.5, ymax = 15.5, xmin = -5.5) +
  annotation_custom(lay, ymin = 13.5, ymax = 15.5, xmin = -3.5) +
  annotation_custom(inc, ymin = 13.5, ymax = 15.5, xmin = -1.5) +
  annotation_custom(inc, ymin = 13.5, ymax = 15.5, xmin = 0.5) +
  annotation_custom(inc, ymin = 13.5, ymax = 15.5, xmin = 2.5) +
  annotation_custom(hatch, ymin = 13.5, ymax = 15.5, xmin = 4.5) +
  annotation_custom(nestling, ymin = 13.5, ymax = 15.5, xmin = 6.5) +
  annotation_custom(nestling, ymin = 13.5, ymax = 15.5, xmin = 8.5), 
          nrow = 2,rel_heights = c(0.4,0.6)) 
```

@@

p3 <- candidates.pit %>% 
  dplyr::filter(treatment %in% charlevels) %>%
  ggplot(aes(x = treatment, y = BRCA1, fill = treatment, color = sex)) +
    geom_boxplot() +
    theme_B3() +
      scale_alpha_manual(values = c(0.75,1)) +
  scale_color_manual(values = sexcolors) +
  scale_fill_manual(values = colorscharnew) +
  theme(legend.position = "none") + 
  labs(x = "increasing time >>", y = "BRCA1", subtitle = "Pituitary") +
  annotation_custom(control, ymin = 6.65, ymax = 7, xmin = -7.8) +
  annotation_custom(bldg, ymin = 6.65, ymax = 7, xmin = -5.8) +
  annotation_custom(lay, ymin = 6.65, ymax = 7, xmin = -3.8) +
  annotation_custom(inc, ymin = 6.65, ymax = 7, xmin = -1.8) +
  annotation_custom(inc, ymin = 6.65, ymax = 7, xmin = 0.6) +
  annotation_custom(inc, ymin = 6.65, ymax = 7, xmin = 2.4) +
  annotation_custom(hatch, ymin = 6.65, ymax = 7, xmin = 4.4) +
  annotation_custom(nestling, ymin = 6.65, ymax = 7, xmin = 6.4) +
  annotation_custom(nestling, ymin = 6.65, ymax = 7, xmin = 8.4) +
  ylim(6.8,8.5) +
  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) +
  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
  theme(axis.title.y  = element_text(face = "italic"))
p3

@@

p2 <- candidates.pit %>%
  dplyr::filter(treatment %in% charlevels) %>%
  ggplot(aes(x = treatment, y = PRL)) + 
    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
    scale_fill_manual(values = colorscharmaip) +
    scale_alpha_manual(values = c(0.75,1)) +
    theme_B3() +
  theme(legend.position = "none") + 
  scale_color_manual(values = sexcolors) +
  labs(y = "Pituitary PRL", x = "parental stage" ) +
  theme(axis.title.y  = element_text(face = "italic"),
        axis.text.x = element_blank())  +
  annotation_custom(control, ymin = 13.5, ymax = 15, xmin = -7.5) +
  annotation_custom(bldg, ymin = 13.5, ymax = 15, xmin = -5.5) +
  annotation_custom(lay, ymin = 13.5, ymax = 15, xmin = -3.5) +
  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = -1.5) +
  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 0.5) +
  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 2.5) +
  annotation_custom(hatch, ymin = 13.5, ymax = 15, xmin = 4.5) +
  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 6.5) +
  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 8.5)

@@

>>>>

source("../R/themes.R") 
source("../R/functions.R")

@@

```{r data}
hormones <- read_csv("../results/07_hormones.csv")
hormones$treatment <- factor(hormones$treatment, levels = alllevels)

@@

@@

## Prolactin expression

```{r candidatevsds}
prl <- read_csv("../results/candidatecounts.csv") %>%
  filter(gene ==  "PRL")  %>% 
  mutate(bird_id = sapply(strsplit(samples, '_'), "[", 1)) %>%
  select(-X1, -samples) %>% 
  select(bird_id, everything())

prl

hormonesslim <- hormones %>%
  select(bird_id, hormone, plasma_conc, logconc)
head(hormonesslim)

hormonesandPRL <- full_join(prl, hormonesslim, by = "bird_id")
hormonesandPRL
```

p2 <- candidates.pit %>%
  dplyr::filter(treatment %in% charlevels) %>%
  ggplot(aes(x = treatment, y = PRL)) + 
    geom_boxplot(aes(fill = treatment, color = sex)) + 
    theme_B3() +
  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
  scale_color_manual(values = sexcolors) +
  scale_fill_manual(values = colorscharnew) +
  labs(y = "PRL", x = "parental care stages", subtitle = "Pituitary" ) +
  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) +
  theme(axis.title.y  = element_text(face = "italic")) 
p2

@@

@@

p3 <- candidates.pit %>% 
  dplyr::filter(treatment %in% charlevels) %>%
  ggplot(aes(x = treatment, y = BRCA1, fill = treatment, color = sex)) +
    geom_boxplot() +
    theme_B3() +
      scale_alpha_manual(values = c(0.75,1)) +
  scale_color_manual(values = sexcolors) +
  scale_fill_manual(values = colorscharnew) +
  theme(legend.position = "none") + 
  labs(x = "increasing time >>", y = "BRCA1", subtitle = "Pituitary") +

  ylim(6.8,8.5) +
  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) +
  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
  theme(axis.title.y  = element_text(face = "italic"))
p3

@@

p2 <- candidates.pit %>%
  dplyr::filter(treatment %in% charlevels) %>%
  ggplot(aes(x = treatment, y = PRL)) + 
    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
    scale_fill_manual(values = colorscharmaip) +
    scale_alpha_manual(values = c(0.75,1)) +
    theme_B3() +
  theme(legend.position = "none") + 
  scale_color_manual(values = sexcolors) +
  labs(y = "Pituitary PRL", x = "parental stage" ) +
  theme(axis.title.y  = element_text(face = "italic"),
        axis.text.x = element_blank())

@@

write.csv(prl, "../results/16_PRL.csv", row.names = F)
write.csv(hormonesandPRL, "../results/16_hormonesandPRL.csv", row.names = F)
